---
title: "R Notebook"
output: html_notebook
---
Exploratory visualizations - What is in the data

Explanatory visualizations - What have we learned from the data


```{r}
plot(ChickWeight)
```
```{r}
library(MASS)
whiteside
```

Checking out the structure of the WhiteSide package
```{r}
str(whiteside)
```

```{r}
plot(whiteside)
```
Creating an exploratory scatterplot
```{r}
plot(whiteside$Temp, whiteside$Gas, xlab='Outside temperature', ylab='Heating gas consumption')
```
Plot is a generic function, it will present the visualization according to the type of data
```{r}
# Apply the plot() function to Insul
plot(whiteside$Insul)
```
Basic visualization of cars 93 dataset
```{r}
# Cars93 is in your workspace
str(Cars93)

# Plot Max.Price vs. Price as red triangles
plot(Cars93$Price, Cars93$Max.Price, col = 'red', pch=17)

# Add Min.Price vs. Price as blue circles
points(Cars93$Price, Cars93$Min.Price, pch=16, col='blue')

# Add an equality reference line with abline()
abline(a = 0, b = 1, lty = 2)
```
Creating multiple plots in the same plot - Using par function
In the below visualization, we have learnt to plot multple graphs in the same image with different titles. Basically everything between each plot statement is attributed to the initial plot.
```{r}
library(robustbase)
# Animals2 is in your workspace
str(Animals2)

# Set up the side-by-side plot array
par(mfrow = c(1,2))
# First plot: brain vs. body in its original form
plot(Animals2$body, Animals2$brain)

# Add the first title
title('Original representation')

# Second plot: log-log plot of brain vs. body
plot(Animals2$body, Animals2$brain, log = 'xy')

# Add the second title
title('Log-log plot')
```
Pie charts - Avoid as much as possible. Can be used for explanatory(to be used judiciously)
```{r}
library(insuranceData)
# dataCar is in your workspace
data("dataCar")
str(dataCar)

# Set up a side-by-side plot array
par(mfrow = c(1, 2))

# Create a table of veh_body record counts and sort
tbl <- sort(table(dataCar$veh_body),
            decreasing = TRUE)

# Create the pie chart
pie(tbl)

# Give it a title
title("Pie chart")

# Create the barplot with perpendicular, half-sized labels
barplot(tbl, las = 2, cex.names = 0.5)

# Add a title
title("Bar chart")
```

Characterising a single variable
Most popular plots: 
Histogram
Scatter
Density
QQ - Heteroskedasticity/Homo - Data points should lie within the 95% confidence interval line

Assumption : All values of numeric should follow gaussian normal distribution. If not, they should be converted to achieve the same. This would be best for our models

----------------------------

hist(uses traditional frequency) vs truehist(uses probability density)
```{r}
# Set up a side-by-side plot array
par(mfrow=c(1,2))

# Create a histogram of counts with hist()
hist(Cars93$Horsepower, main='hist() plot')

# Create a normalized histogram with truehist()
truehist(Cars93$Horsepower, main='truehist() plot')
```

Density plots as smoothened histograms
Adding density curve line
Chickwt dataset from datasets package

```{r}
# Create index16, pointing to 16-week chicks
index16 <- which(ChickWeight$Time == 16)

# Get the 16-week chick weights
weights <- ChickWeight$weight[index16]

# Plot the normalized histogram
truehist(weights)

# Add the density curve to the histogram
lines(density(weights))
```

Real data => Difficult to determine probability distribution
Enter QQ Plot - Will tell us exactly if the data follows gaussian distribution or not
```{r}
# Load the car package to make qqPlot() available
library(car)

# Create index16, pointing to 16-week chicks
index16 <- which(ChickWeight$Time == 16)

# Get the 16-week chick weights
weights <- ChickWeight$weight[index16]

# Show the normal QQ-plot of the chick weights
qqPlot(weights)

# Show the normal QQ-plot of the Boston$tax data
qqPlot(Boston$tax)
```

---------------------

Visualizing relations between variables

X and y are numeric

Scatterplots - plot function(x,y) or y~x, data=Datasource

Sunflower function - Represent repeated data points as sunflowers - Each iteration is each 

X is numeric , Y is categorical
Boxplots


Mosaic plots - X and Y are categorical or X and Y are integer with multiple repeated valeus

--------------------
Scatterplots - Used to 

```{r}
# Set up a side-by-side plot array
par(mfrow=c(1,2))

# Create the standard scatterplot
plot(Boston$zn, Boston$rad)

# Add the title
title('Standard scatterplot')

# Create the sunflowerplot
sunflowerplot(Boston$zn, Boston$rad)

# Add the title
title('Sunflower plot')
```

Boxplot - Shows how value of variable differ with respect to unique levels of secondary variable(<10 unique levels ~ good output)
```{r}
# Create a variable-width boxplot with log y-axis & horizontal labels
boxplot(crim ~ rad, data = Boston, varwidth=T, log='y', las=1 )

# Add a title
title('Crime rate vs. radial highway index')
```

Mosiac plot - Scatter plot between categorical variables

```{r}
# Create a mosaic plot using the formula interface
mosaicplot(carb ~ cyl, data = mtcars)
```

Representing more complex relationships between variables

Bagplot - 2D Boxplot joint distribution of two variables
Red star - Median
Dark blue - Box - 50% datapoints
Light blue - Fence 
Pts outside are outliers

Correlation coefficient - Useful preliminary view
Relationship between two numerical varibales
-1 < r < 1
cor - correlation matrix
---- Similar to rattle

rpart - Decision tree models
rpart - used to fit the tree model
text - adds labels to the nodes

```{r}
# Load aplpack to make the bagplot() function available
library(aplpack)

# Create a side-by-side boxplot summary
par(mfrow=c(1,2))
boxplot(Cars93$Min.Price, Cars93$Max.Price)

# Create a bagplot for the same two variables
bagplot(Cars93$Min.Price, Cars93$Max.Price, cex=1.2)

# Add an equality reference line
abline(0,1, lty=2)
```

Correlation matrix----------------


```{r}
# Load the corrplot library for the corrplot() function
library(corrplot)

# Extract the numerical variables from UScereal
numericalVars <- UScereal[,unlist(lapply(UScereal, is.numeric))]

# Compute the correlation matrix for these variables
corrMat <- cor(numericalVars)

# Generate the correlation ellipse plot
corrplot(corrMat, method='ellipse')
```
Plotting decision based models

```{r}
# Load the rpart library
library(rpart)

# Fit an rpart model to predict medv from all other Boston variables
tree_model <- rpart(medv ~., data= Boston)

# Plot the structure of this decision tree model
plot(tree_model)

# Add labels to this plot
text(tree_model, cex=0.7)
```


Plot()
global function - par - params : mfrow, cex.main (Specifies the size)
pch - shape of points

high level - plots(eg. plot())
low level - adding features to plots
lines, titles

```{r}
## Exploring the par function and it's parameters
# Assign the return value from the par() function to plot_pars
plot_pars <- par()

# Display the names of the par() function's list elements
names(plot_pars)

# Display the number of par() function list elements
length(plot_pars)
```

```{r}
# Set up a 2-by-2 plot array
par(mfrow=c(2,2))

# Plot the Animals2 brain weight data as points
plot(Animals2$brain, type = "p")

# Add the title
title("points")

# Plot the brain weights with lines
plot(Animals2$brain, type = "l")

# Add the title
title("lines")

# Plot the brain weights as lines overlaid with points
plot(Animals2$brain, type = "o")

# Add the title
title("overlaid")

# Plot the brain weights as steps
plot(Animals2$brain, type = "s")

# Add the title
title("steps")
```
Plotting multiple data frames on the same plot - Exploring the pch types :)
```{r}
# Compute max_hp
max_hp <- max(Cars93$Horsepower, mtcars$hp)

# Compute max_mpg
max_mpg <- max(Cars93$MPG.city, Cars93$MPG.highway, mtcars$mpg)

# Create plot with type = "n"               
plot(max_hp, max_mpg,
     type = "n", xlim = c(0, max_hp),
     ylim = c(0, max_mpg), xlab = "Horsepower",
     ylab = "Miles per gallon")

# Add open circles to plot
points(mtcars$hp, mtcars$mpg, pch = 1)

# Add solid squares to plot
points(Cars93$Horsepower, Cars93$MPG.city, pch = 15)

# Add open triangles to plot
points(Cars93$Horsepower, Cars93$MPG.highway, pch = 2)
```

Hist - probability
truhist - normalized probability
lines - adding lines to a plot(x,y)
abline - adds straight lines (slope, intercept)

```{r}
## Exploring the lines function
# Create the numerical vector x
x <- seq(0, 10, length = 200)

# Compute the Gaussian density for x with mean 2 and standard deviation 0.2
gauss1 <- dnorm(x, mean = 2, sd = 0.2)

# Compute the Gaussian density with mean 4 and standard deviation 0.5
gauss2 <- dnorm(x, mean = 4, sd = 0.5)

# Plot the first Gaussian density
plot(x, gauss1, ylab="Gaussian probability density", type="l")

# Add lines for the second Gaussian density
lines(x, gauss2, lty=2, lwd=3)
```

```{r}
## Exploring point options
# Create an empty plot using type = "n"
plot(mtcars$hp, mtcars$mpg, type="n", xlab="Horsepower", ylab="Gas mileage")

# Add points with shapes determined by cylinder number

points(mtcars$hp, mtcars$mpg, pch=mtcars$cyl)

# Create a second empty plot
plot(mtcars$hp, mtcars$mpg, type="n", xlab="Horsepower", ylab="Gas mileage")

# Add points with shapes as cylinder characters
points(mtcars$hp, mtcars$mpg, pch = as.character(mtcars$cyl))
```

Creating a regression model and adding reference line
```{r}
# Build a linear regression model for the whiteside data
linear_model <- lm(Gas ~ Temp, data=whiteside)

# Create a Gas vs. Temp scatterplot from the whiteside data
plot(whiteside$Temp, whiteside$Gas)

# Use abline() to add the linear regression line
abline(linear_model, lty=2)
```

Adding text to plot 
Plot + Text = Great visualization
axis labels
titles
legends
text in plots \n => new line
text(x,y labels, adj - 0.5 0(Left alignment, 1 Right aligned))
srt= angle of text orientation with respect to x axis
col - color of the text can be altered using this parameter
cex=1.2 - 20% larger
font -1 normal 2 bold 3 italics 4 bold and italics

```{r}

## Exploring text within plots
# Create MPG.city vs. Horsepower plot with solid squares
plot(Cars93$Horsepower, Cars93$MPG.city, pch=15)

# Create index3, pointing to 3-cylinder cars
index3<-which(Cars93$Cylinders == 3)

# Add text giving names of cars next to data points
text(x = Cars93$Horsepower[index3], 
     y = Cars93$MPG.city[index3],
     labels = Cars93$Make[index3], adj = 0)
```

```{r}

### Some more text properties
# Plot MPG.city vs. Horsepower as open circles
plot(Cars93$Horsepower, Cars93$MPG.city, pch=1)

# Create index3, pointing to 3-cylinder cars
index3 <- which(Cars93$Cylinders ==3)

# Highlight 3-cylinder cars as solid circles
points(Cars93$Horsepower[index3], Cars93$MPG.city[index3], pch=16)

# Add car names, offset from points, with larger bold text
text(Cars93$Horsepower[index3], Cars93$MPG.city[index3], labels=Cars93$Make[index3], adj = -0.2, cex=1.2, font=4 )
```

```{r}
### Playing around with labels god mode
# Plot Gas vs. Temp as solid triangles
plot(whiteside$Temp, whiteside$Gas, pch=17)

# Create indexB, pointing to "Before" data
indexB<- which(whiteside$Insul == "Before")

# Create indexA, pointing to "After" data
indexA<- which(whiteside$Insul == "After")

# Add "Before" text in blue, rotated 30 degrees, 80% size
text(x = whiteside$Temp[indexB], y =whiteside$Gas[indexB],
     labels = "Before", col = "blue", srt = 30, cex = 0.8)

# Add "After" text in red, rotated -20 degrees, 80% size
text(x = whiteside$Temp[indexA], y =whiteside$Gas[indexA],
     labels ="After", col = "red", srt = -20, cex = 0.8)

```

Legends - X legend position, pch and col can be added easily

```{r}
# Set up and label empty plot of Gas vs. Temp
plot(whiteside$Temp, whiteside$Gas,
     type = "n", xlab = "Outside temperature",
     ylab = "Heating gas consumption")

# Create indexB, pointing to "Before" data
indexB <- which(whiteside$Insul == "Before")

# Create indexA, pointing to "After" data
indexA <- which (whiteside$Insul == "After")

# Add "Before" data as solid triangles
points(whiteside$Temp[indexB], whiteside$Gas[indexB], pch=17)

# Add "After" data as open circles
points(whiteside$Temp[indexB], whiteside$Gas[indexB], pch = 1)

# Add legend that identifies points as "Before" and "After"
legend("topright", pch = c(17,1), 
       legend = c("Before", "After"))
```

```{r}
# Create a boxplot of sugars by shelf value, without axes
boxplot(sugars ~ shelf, data = UScereal,
        axes = FALSE)

# Add a default y-axis to the left of the boxplot
axis(side = 2)

# Add an x-axis below the plot, labelled 1, 2, and 3
axis(side = 1, at = c(1, 2, 3))

# Add a second x-axis above the plot
axis(side = 3, at = c(1, 2, 3),
     labels = c("floor", "middle", "top"))
```

```{r}
### Playing around with trend lines
# Create a scatterplot of MPG.city vs. Horsepower
plot(Cars93$Horsepower, Cars93$MPG.city)

# Call supsmu() to generate a smooth trend curve, with default bass
trend1 <- supsmu(Cars93$Horsepower, Cars93$MPG.city, bass=0)

# Add this trend curve to the plot
lines(trend1)

# Call supsmu() for a second trend curve, with bass = 10
trend2 <- supsmu(Cars93$Horsepower, Cars93$MPG.city, bass=10)

# Add this trend curve as a heavy, dotted line
lines(trend2, lty=3, lwd=2)

```
Managing visual complexity
Too small - Visual utility loss, too much information on the plots
Too large - Not visually appealing

Alt:
Multiple plots on the same axis using differential techniques color encoding, legends

How much detail is too much?

```{r}
## Too much detail, too many plots
# Compute the number of plots to be displayed
ncol(Cars93)^2

# Plot the array of scatterplots
plot(Cars93)
```


Using the matplot function to create multiple plots on the same plot
```{r}
# Construct the vector keep_vars
keep_vars <- c("calories", "protein", "fat",
               "fibre", "carbo", "sugars")

# Use keep_vars to extract the desired subset of UScereal
df <- UScereal[, keep_vars]

# Set up a two-by-two plot array
par(mfrow = c(2,2))

# Use matplot() to generate an array of two scatterplots
matplot(df$calories, df[, c("protein","fat")], xlab = "calories", ylab="")
# Add a title
title("Two scatterplots")

# Use matplot() to generate an array of three scatterplots
matplot(df$calories, df[, c("protein","fat","fibre")], xlab = "calories", ylab="")
# Add a title
title("Three scatterplots")

# Use matplot() to generate an array of four scatterplots
matplot(df$calories, df[, c("protein","fat","fibre","carbo")], xlab = "calories", ylab="")

# Add a title
title("Four scatterplots")

# Use matplot() to generate an array of five scatterplots
matplot(df$calories, df[, c("protein","fat","fibre","carbo","sugars")], xlab = "calories", ylab="")

# Add a title
title("Five scatterplots")
```
Wordcloud package
Min freq value =  3 by default, can be changed to 1 for all manufacturers
```{r}
library(wordcloud)
# Create mfr_table of manufacturer frequencies
mfr_table <- table(Cars93$Manufacturer)
head(mfr_table)
# Create the default wordcloud from this table
wordcloud(words = names(mfr_table), 
          freq = as.numeric(mfr_table), 
          scale = c(2,0.25))

# Change the minimum word frequency
wordcloud(words = names(mfr_table), 
          freq = as.numeric(mfr_table), 
          scale = c(2,0.25), 
          min.freq = 1)

# Create model_table of model frequencies
model_table <- table(Cars93$Model)

# Create the wordcloud of all model names with smaller scaling
wordcloud(words = names(model_table), 
          freq = as.numeric(model_table), 
          scale = c(0.75,0.25), 
          min.freq = 1)
```

Creating plot arrays with mfrow argument

Multiple plot arrays - Differences/different views of dataset

Pairs function(sets up two columns)

mfrow - multiple plots
layout - multiple plots

plot size
pty = s(square), im (Max available space)

Creating multiple plots in the same window
```{r}
# Set up a two-by-two plot array
par(mfrow = c(2,2))

# Plot y1 vs. x1 
plot(anscombe$x1, anscombe$y1)

# Plot y2 vs. x2
plot(anscombe$x2, anscombe$y2)

# Plot y3 vs. x3
plot(anscombe$x3, anscombe$y3)

# Plot y4 vs. x4
plot(anscombe$x4, anscombe$y4)
```
Common Scaling - Is used to set the common limites of each variable. This means that different datasets can be compared easily since juxtaposing datasets on same axis can serve as preemptive measure
```{r}
# Define common x and y limits for the four plots
xmin <- min(anscombe$x1,anscombe$x2,anscombe$x3,anscombe$x4)
xmax <- max(anscombe$x1,anscombe$x2,anscombe$x3,anscombe$x4)
ymin <- round(min(anscombe$y1,anscombe$y2,anscombe$y3,anscombe$y4))
ymax <- round(max(anscombe$y1,anscombe$y2,anscombe$y3,anscombe$y4))

# Set up a two-by-two plot array
par(mfrow=c(2,2))

# Plot y1 vs. x1 with common x and y limits, labels & title
plot(anscombe$x1, anscombe$y1,
     xlim = c(xmin, xmax),
     ylim = c(ymin, ymax),
     xlab = "x value", ylab = "y value",
     main = "First dataset")

# Do the same for the y2 vs. x2 plot
plot(anscombe$x2, anscombe$y2,
     xlim = c(xmin, xmax),
     ylim = c(ymin, ymax),
     xlab = "x value", ylab = "y value",
     main = "Second dataset")

# Do the same for the y3 vs. x3 plot
plot(anscombe$x3, anscombe$y3,
     xlim = c(xmin, xmax),
     ylim = c(ymin, ymax),
     xlab = "x value", ylab = "y value",
     main = "Third dataset")

# Do the same for the y4 vs. x4 plot
plot(anscombe$x4, anscombe$y4,
     xlim = c(xmin, xmax),
     ylim = c(ymin, ymax),
     xlab = "x value", ylab = "y value",
     main = "Fourth dataset")
```

Exploring different data graphs plotted on the same axis
```{r}
# Set up a two-by-two plot array
par(mfrow=c(2,2))

# Plot the raw duration data
plot(geyser$duration, main='Raw data')

# Plot the normalized histogram of the duration data
truehist(geyser$duration, main='Histogram')

# Plot the density of the duration data
plot(density(geyser$duration), main='Density')

# Construct the normal QQ-plot of the duration data
qqPlot(geyser$duration, main="QQ-plot")
```
Creating plots with layout function()
layout - uses matrix to define the plot array
```{r}
## Playing around with layout function
# Use the matrix function to create a matrix with three rows and two columns
layoutMatrix <- matrix(
  c(
    0, 1,
    2, 0,
    0, 3
  ), 
  byrow = T, 
  nrow = 3
)

# Call the layout() function to set up the plot array
layout(layoutMatrix)

# Show where the three plots will go 
layout.show(n=3)
```
Advanced multiple plotting using layout function
```{r}
# Set up the plot array
layout(layoutMatrix)

# Construct vectors indexB and indexA
indexB <- which(whiteside$Insul == "Before")
indexA <- which(whiteside$Insul =="After")

# Create plot 1 and add title
plot(whiteside$Temp[indexB], whiteside$Gas[indexB],
     ylim = c(0,8))
title("Before data only")

# Create plot 2 and add title
plot(whiteside$Temp, whiteside$Gas,
     ylim = c(0,8))
title("Complete dataset")



# Create plot 3 and add title
plot(whiteside$Temp[indexA], whiteside$Gas[indexA],
     ylim = c(0,8))
title("After data only")
```
Creating different sized subplots - luxury with layout function - cannot be done using par function
```{r}
# Create row1, row2, and layoutVector
row1 <- c(1,0,0)
row2 <- c(0,2,2)
layoutVector <- c(row1,row2, row2)

# Convert layoutVector into layoutMatrix
layoutMatrix <- matrix(layoutVector, byrow = T, nrow = 3)

# Set up the plot array
layout(layoutMatrix)

# Plot scatterplot
plot(Boston$rad, Boston$zn)

# Plot sunflower plot
sunflowerplot(Boston$rad, Boston$zn)
```
Creating and saving more complex plots

----------

Side effects - Generate plots but no return values
barplot() - returns numerical value
scatterplots()

symbols() - data points can be customized using this

bg = color

plots can be saved as image files
png()

```{r}
## Adding text to labels
# Create a table of Cylinders frequencies
tbl <- table(Cars93$Cylinders)

# Generate a horizontal barplot of these frequencies
mids <- barplot(tbl, horiz = T,
                col ="transparent",
                names.arg = "")

# Add names labels with text()
text(20, mids, labels=names(tbl))

# Add count labels with text()
text(35, mids, labels = as.numeric(tbl))
```

```{r}
## Exploring the symbols function
# Call symbols() to create the default bubbleplot
symbols(Cars93$Horsepower, Cars93$MPG.city,
        circles = sqrt(Cars93$Price))

# Repeat, with the inches argument specified
symbols(Cars93$Horsepower, Cars93$MPG.city,
        circles = sqrt(Cars93$Price),
        inches = 0.1)
```
Saving plot as png file:
```{r}
# Call png() with the name of the file we want to create
png("bubbleplot.png")

# Re-create the plot from the last exercise
symbols(Cars93$Horsepower, Cars93$MPG.city,
        circles = sqrt(Cars93$Price),
        inches = 0.1)

# Save our file and return to our interactive session
dev.off()

# Verify that we have created the file
list.files(pattern = "png")
```
Using color effectively

Different colors for differnt categorical variables
Limitations:
-Color blind
-Redistributed as black and white
-Too many colors, too less colors 

Sol:
Ideally 6
Max 12
No more than 20

Adding color to life ;)
```{r}
# Iliinsky and Steele color name vector
IScolors <- c("red", "green", "yellow", "blue",
              "black", "white", "pink", "cyan",
              "gray", "orange", "brown", "purple")

# Create the data for the barplot
barWidths <- c(rep(2, 6), rep(1, 6))

# Recreate the horizontal barplot with colored bars
barplot(rev(barWidths), horiz = T,
        col = rev(IScolors), axes = F,
        names.arg = rev(IScolors), las = 1)
```
Advanced coloring options
```{r}
# Iliinsky and Steele color name vector
IScolors <- c("red", "green", "yellow", "blue",
              "black", "white", "pink", "cyan",
              "gray", "orange", "brown", "purple")

# Create the `cylinderLevel` variable
cylinderLevels <- as.numeric(Cars93$Cylinders)

# Create the colored bubbleplot
symbols(Cars93$Horsepower, Cars93$MPG.city, 
        circles = cylinderLevels, inches = 0.2, 
        bg = IScolors[cylinderLevels])
```
```{r}
### Creating a stacked bar graph

# Create a table of Cylinders by Origin
tbl <- table(Cars93$Cylinders, Cars93$Origin)

# Create the default stacked barplot
barplot(tbl)

# Enhance this plot with color
barplot(tbl, col=IScolors)
```

Graphic systems in R
1. Base - Good for exploratory analysis
2. Grid - grid package - greater control over low level graphical details
grid-base - hybrid of base and grid
3. Lattice - extension of grid graphics - eg conditional scatterplot
4. ggplot2 - extension of grid graphics - better for high level visualizations
stages
These are incompatible with each other
```{r}
# Load the insuranceData package
library(insuranceData)

# Use the data() function to load the dataCar data frame
data(dataCar)

# Load the tabplot package
suppressPackageStartupMessages(library(tabplot))

# Generate the default tableplot() display
tableplot(dataCar)
```

Enter ggplot
```{r}
# Load the ggplot2 package
library(ggplot2)

# Create the basic plot (not displayed): basePlot
basePlot <- ggplot(Cars93, aes(x = Horsepower, y = MPG.city))

# Display the basic scatterplot
basePlot + 
  geom_point()

# Color the points by Cylinders value
basePlot + 
  geom_point(colour = IScolors[Cars93$Cylinders])

# Make the point sizes also vary with Cylinders value
basePlot + 
  geom_point(colour = IScolors[Cars93$Cylinders], 
             size = as.numeric(Cars93$Cylinders))
```


