---
title: "GGPlot part 1"
output: html_notebook
---
Visualization using ggplot2

viz = Stats + design

Exploratory - Specific audience
Explanatory - Broader audience

Using color, different datasets, understand common pitfalls and best practices

Basic plot
```{r}
# Load the ggplot2 package
library(ggplot2)

# Explore the mtcars data frame with str()
str(mtcars)

# Execute the following command
ggplot(mtcars, aes(x = cyl, y = mpg)) +
  geom_point()
```

Dealing with categorical variables
```{r}
# Load the ggplot2 package
library(ggplot2)

# Change the command below so that cyl is treated as factor
ggplot(mtcars, aes(x = factor(cyl), y = mpg)) +
  geom_point()
```

Grammar of graphics
Plotting framework
- Graphics = distinct layers of grammatical elements
- Meaningful plots through aesthetic mapping

7 grammatical elements
Data
Aesthetics - axis/scale
Geometrics - visual elements for data
Facets - small multiples
Statistics - Understanding statistical significance of data
Coordinates - Space on which the data is plotted
Themes - Non data ink
```{r}
### Exploring the basic ggplot features

# A scatter plot has been made for you
ggplot(mtcars, aes(x = wt, y = mpg)) +
  geom_point()

# Replace ___ with the correct column
ggplot(mtcars, aes(x = wt, y = mpg, color = disp)) +
  geom_point()

# Replace ___ with the correct column
ggplot(mtcars, aes(x = wt, y = mpg, size = disp)) +
  geom_point()

```
Diamonds dataset
```{r}
str(diamonds)

# Add geom_point() with +
ggplot(diamonds, aes(x = carat, y = price)) + geom_point()

# Add geom_point() and geom_smooth() with +
ggplot(diamonds, aes(x = carat, y = price))+geom_point()+geom_smooth()
```
Exploring ggplot2 plotting
```{r}
# 1 - The plot you created in the previous exercise
ggplot(diamonds, aes(x = carat, y = price)) +
  geom_point() +
  geom_smooth()

# 2 - Copy the above command but show only the smooth line
ggplot(diamonds, aes(x = carat, y = price)) +
  geom_smooth()


# 3 - Copy the above command and assign the correct value to col in aes()
ggplot(diamonds, aes(x = carat, y = price, colour=clarity)) +
  geom_smooth()


# 4 - Keep the color settings from previous command. Plot only the points with argument alpha.
ggplot(diamonds, aes(x = carat, y = price, colour=clarity)) +
  geom_point(alpha=0.4)

```
```{r}
# Create the object containing the data and aes layers: dia_plot
dia_plot <- ggplot(diamonds, aes(x = carat, y = price))

# Add a geom layer with + and geom_point()
dia_plot + geom_point()

# Add the same geom layer, but with aes() inside
dia_plot + geom_point(aes(color = clarity))

```
Adding additional smoothening functions:
```{r}
# 1 - The dia_plot object has been created for you
dia_plot <- ggplot(diamonds, aes(x = carat, y = price))

# 2 - Expand dia_plot by adding geom_point() with alpha set to 0.2
dia_plot + geom_point(alpha=0.2)

# 3 - Plot dia_plot with additional geom_smooth() with se set to FALSE
dia_plot + geom_smooth(se = F)

# 4 - Copy the command from above and add aes() with the correct mapping to geom_smooth()
dia_plot + geom_smooth(aes(col = clarity), se = F)

```
Data layer
Base plot() disadvantages
- Plot is an image
- Plot does not get redrawn
- Legend needs to be added manually
- No unified framework for plotting

ggplot() - It is an object
More flexibility
Reusagble data
```{r}
# Plot the correct variables of mtcars
plot(mtcars$wt, mtcars$mpg, col = mtcars$cyl)

# Change cyl inside mtcars to a factor
mtcars$fcyl <- as.factor(mtcars$cyl)

# Make the same plot as in the first instruction
plot(mtcars$wt, mtcars$mpg, col=mtcars$fcyl)
```
Plot function linear model
Key takeaways:
Using base package plot function, we can add the linear model, however, it is a tedious task since we have to add legend manually
```{r}
# Use lm() to calculate a linear model and save it as carModel
carModel <- lm(mpg ~ wt, data = mtcars)

# Basic plot
mtcars$cyl <- as.factor(mtcars$cyl)
plot(mtcars$wt, mtcars$mpg, col = mtcars$cyl)

# Call abline() with carModel as first argument and set lty to 2
abline(carModel, lty = 2)

# Plot each subset efficiently with lapply
# You don't have to edit this code
## Basically - Three different ablines are added with respect to cyl value
plot(mtcars$wt, mtcars$mpg, col = mtcars$cyl)
lapply(mtcars$cyl, function(x) {
  abline(lm(mpg ~ wt, mtcars, subset = (cyl == x)), col = x)
  })

# This code will draw the legend of the plot
# You don't have to edit this code
legend(x = 5, y = 33, legend = levels(mtcars$cyl),
       col = 1:3, pch = 1, bty = "n")
```
```{r}
# Convert cyl to factor (don't need to change)
mtcars$cyl <- as.factor(mtcars$cyl)

# Example from base R (don't need to change)
plot(mtcars$wt, mtcars$mpg, col = mtcars$cyl)
abline(lm(mpg ~ wt, data = mtcars), lty = 2)
lapply(mtcars$cyl, function(x) {
  abline(lm(mpg ~ wt, mtcars, subset = (cyl == x)), col = x)
  })
legend(x = 5, y = 33, legend = levels(mtcars$cyl),
       col = 1:3, pch = 1, bty = "n")

# Plot 1: add geom_point() to this command to create a scatter plot
ggplot(mtcars, aes(x = wt, y = mpg, col = cyl)) +
  geom_point()  # Fill in using instructions Plot 1

# Plot 2: include the lines of the linear models, per cyl
ggplot(mtcars, aes(x = wt, y = mpg, col = cyl)) +
  geom_point() + # Copy from Plot 1
  geom_smooth(method='lm', se=F)   # Fill in using instructions Plot 2

# Plot 3: include a lm for the entire dataset in its whole
ggplot(mtcars, aes(x = wt, y = mpg, col = cyl)) +
  geom_point() + # Copy from Plot 2
  geom_smooth(method = 'lm', se=F) + # Copy from Plot 2
  geom_smooth(aes(group=1), method='lm', se=F, linetype=2)  # Fill in using instructions Plot 3
```
ggplot
Tidy data - Each column is a single variable
Use small multiples to juxtapose plots
```{r}
# Consider the structure of iris, iris.wide and iris.tidy (in that order)
str(iris)

# Load the tidyr package
library(tidyr)

# Fill in the ___ to produce to the correct iris.tidy dataset
iris.tidy <- iris %>%
  gather(key, Value, -Species) %>%
  separate(key, c("Part", "Measure"), "\\.")

str(iris.tidy)


# Think about which dataset you would use to get the plot shown right
# Fill in the ___ to produce the plot given to the right
ggplot(iris.tidy, aes(x = Species, y = Value, col = Part)) +
  geom_jitter() +
  facet_grid(. ~ Measure)
```

```{r}
# The 3 data frames (iris, iris.wide and iris.tidy) are available in your environment
# Execute head() on iris, iris.wide and iris.tidy (in that order)
head(iris)

head(iris.tidy)

# Load the tidyr package
library(tidyr)

# Add column with unique ids (don't need to change)
iris$Flower <- 1:nrow(iris)

# Fill in the ___ to produce to the correct iris.wide dataset
iris.wide <- iris %>%
  gather(key, value, -Species, -Flower) %>%
  separate(key, c("Part", "Measure"), "\\.") %>%
  spread(Measure, value)

# Think about which dataset you would use to get the plot shown right
# Fill in the ___ to produce the plot given to the right
ggplot(iris.wide, aes(x = Length, y = Width, color = Part)) +
  geom_jitter() +
  facet_grid(. ~ Species)
```

Aesthetics
  X and Y axis mappings
  aes - aes def
  attributes - geom layer
  col
  fill - fill colour
  size 
  alpha
  linetype - has pattern
  labels - text on plot or axes
  
```{r}
## Playing around with aesthetics
# 1 - Map mpg to x and cyl to y
ggplot(mtcars, aes(mpg, cyl)) +
  geom_point()
  
# 2 - Reverse: Map cyl to x and mpg to y
ggplot(mtcars, aes(cyl, mpg)) +
  geom_point()

# 3 - Map wt to x, mpg to y and cyl to col
ggplot(mtcars, aes(wt, mpg, col=cyl)) +
  geom_point()

# 4 - Change shape and size of the points in the above plot
ggplot(mtcars, aes(wt, mpg, col=cyl)) +
  geom_point(shape=1,size=4)

```
```{r}
# am and cyl are factors, wt is numeric
class(mtcars$am)
class(mtcars$cyl)
class(mtcars$wt)

# Given from the previous exercise
ggplot(mtcars, aes(x = wt, y = mpg, col = cyl)) +
  geom_point(shape = 1, size = 4)

# 1 - Map cyl to fill
ggplot(mtcars, aes(x = wt, y = mpg, fill = cyl)) +
  geom_point(shape = 1, size = 4)

# 2 - Change shape and alpha of the points in the above plot
ggplot(mtcars, aes(x = wt, y = mpg, fill = cyl)) +
  geom_point(shape = 21, size = 4, alpha = 0.6)
  
# 3 - Map am to col in the above plot
ggplot(mtcars, aes(x = wt, y = mpg, fill = cyl, col = am)) +
  geom_point(shape = 21, size = 4, alpha = 0.6)
```

```{r}
# Map cyl to size
ggplot(mtcars, aes(wt,mpg,size=cyl))+ geom_point()

# Map cyl to alpha
ggplot(mtcars, aes(wt,mpg,alpha=cyl))+ geom_point()

# Map cyl to shape 
ggplot(mtcars, aes(wt,mpg,shape=cyl))+ geom_point()


# Map cyl to label
ggplot(mtcars, aes(wt,mpg,label=cyl))+geom_text()


```
Advacned geaom properties
```{r}
# Define a hexadecimal color
my_color <- "#4ABEFF"

# 1 - First scatter plot, with col aesthetic:
ggplot(mtcars, aes(wt, mpg, col=cyl)) + geom_point()


# 2 - Plot 1, but set col attributes in geom layer:
ggplot(mtcars, aes(wt, mpg, col=cyl)) + geom_point(col=my_color)


# 3 - Plot 2, with fill instead of col aesthetic, plut shape and size attributes in geom layer.
ggplot(mtcars, aes(wt, mpg, fill=cyl)) + geom_point(size=10, shape= 23, col=my_color)


```
Playing with color and label
```{r}
# Expand to draw points with alpha 0.5
ggplot(mtcars, aes(x = wt, y = mpg, fill = cyl))+geom_point(alpha=0.5)

  
# Expand to draw points with shape 24 and color yellow
ggplot(mtcars, aes(x = wt, y = mpg, fill = cyl)) + geom_point(shape=24, colour="yellow")

  
# Expand to draw text with label rownames(mtcars) and color red
ggplot(mtcars, aes(x = wt, y = mpg, fill = cyl)) +
  geom_text(label = rownames(mtcars), color = 'red')
```
Advanced graphics using ggplot
```{r}
# Map mpg onto x, qsec onto y and factor(cyl) onto col
ggplot(mtcars, aes(mpg, qsec, col=factor(cyl)))+geom_point()


# Add mapping: factor(am) onto shape
ggplot(mtcars, aes(mpg, qsec,shape=factor(am), col=factor(cyl)))+ geom_point()

# Add mapping: (hp/wt) onto size
ggplot(mtcars, aes(mpg, qsec,size=(hp/wt), shape=factor(am), col=factor(cyl)))+ geom_point()

```
Modifying aesthetics

Position - 
identity - value in dataframe witll pe plotted like data
jittering - adding random noise to identify density
position_jitter(width=0.1)
scale_ - check and modify the scales of plot

scale_ args - name, limit, breaks, expand, labels
labs()

```{r}
# The base layer, cyl.am, is available for you
cyl.am <- ggplot(mtcars, aes(x = factor(cyl), fill = factor(am)))

# Add geom (position = "stack" by default)
cyl.am + 
  geom_bar()

# Fill - show proportion
cyl.am + 
  geom_bar(position = "fill")  

# Dodging - principles of similarity and proximity
cyl.am +
  geom_bar(position ="dodge") 

# Clean up the axes with scale_ functions
val = c("#E41A1C", "#377EB8")
lab = c("Manual", "Automatic")
cyl.am +
  geom_bar(position = "dodge") +
  scale_x_discrete("Cylinders") + 
  scale_y_continuous("Number") +
  scale_fill_manual("Transmission", 
                    values = val,
                    labels = lab) 
```
Adding scales to make the plot appropriately sized
```{r}
# 1 - Create jittered plot of mtcars, mpg onto x, 0 onto y
ggplot(mtcars, aes(x = mpg, y = 0)) +
  geom_jitter()

# 2 - Add function to change y axis limits
ggplot(mtcars, aes(x = mpg, y = 0)) +
  geom_jitter()+
  scale_y_continuous(limits=c(-2,2))

```

Aesthetic best practices
Function - important to identify audience, plot accordingly, data should be represented accurately in the plot
Form - visual aesthetics

Continuous variables - scatter plots (position)
color - not ideal for continuous variables
size - not ideal

Aligned scales - very important(juxtaposed on same axis)

Categorical variables
Color/Size - use should not lead to confusion or excess of information

```{r}
# Basic scatter plot: wt on x-axis and mpg on y-axis; map cyl to col
cyl_fac<-as.factor(mtcars$cyl)
ggplot(mtcars, aes(wt, mpg, col=cyl_fac))+geom_point(size=4)


# Hollow circles - an improvement
ggplot(mtcars, aes(wt, mpg, col=cyl_fac))+geom_point(shape=1, size=4)


# Add transparency - very nice
ggplot(mtcars, aes(wt, mpg, col=cyl_fac))+geom_point(alpha=0.6, size=4)

```
```{r}
# Scatter plot: carat (x), price (y), clarity (color)
ggplot(diamonds, aes(carat, price, col=clarity))+geom_point()

# Adjust for overplotting
ggplot(diamonds, aes(carat, price, col=clarity))+geom_point(alpha=0.5)


# Scatter plot: clarity (x), carat (y), price (color)
ggplot(diamonds, aes(clarity, carat, col=price))+geom_point(alpha=0.5)


# Dot plot with jittering
ggplot(diamonds, aes(clarity, carat, col=price))+geom_point(alpha=0.5, position="jitter")

```
Common plots
 scatter plots
 bar plots
 line plots
 
scatter plots - geom_point(aes(x,y))
overplotting solutions - add jitter, add hollow circles

```{r}
# Shown in the viewer:
ggplot(mtcars, aes(x = cyl, y = wt)) +
  geom_point()

# Solutions:
# 1 - With geom_jitter()
ggplot(mtcars, aes(x = cyl, y = wt)) +
  geom_jitter()

# 2 - Set width in geom_jitter()
ggplot(mtcars, aes(x = cyl, y = wt)) +
  geom_jitter(width=0.1)

# 3 - Set position = position_jitter() in geom_point() ()
ggplot(mtcars, aes(x = cyl, y = wt)) +
  geom_point(position = position_jitter(0.1))

```
```{r}
Vocab<- read.csv("Vocab.csv")

# Examine the structure of Vocab
str(Vocab)

# Basic scatter plot of vocabulary (y) against education (x). Use geom_point()
ggplot(Vocab, aes(education,vocabulary))+geom_point()


# Use geom_jitter() instead of geom_point()
ggplot(Vocab, aes(education,vocabulary))+geom_jitter()



# Using the above plotting command, set alpha to a very low 0.2
ggplot(Vocab, aes(education,vocabulary))+geom_jitter(alpha=0.2)


# Using the above plotting command, set the shape to 1
ggplot(Vocab, aes(education,vocabulary))+geom_jitter(alpha=0.2, shape=1)

```
Enter barplots:
```{r}
# 1 - Make a univariate histogram
ggplot(mtcars, aes(x = mpg)) +
  geom_histogram()

# 2 - Plot 1, plus set binwidth to 1 in the geom layer
ggplot(mtcars, aes(x = mpg)) +
  geom_histogram(binwidth=1)


# 3 - Plot 2, plus MAP ..density.. to the y aesthetic (i.e. in a second aes() function)
ggplot(mtcars, aes(x = mpg)) +
  geom_histogram(binwidth=1, aes(y=..density..))


# 4 - plot 3, plus SET the fill attribute to "#377EB8"
ggplot(mtcars, aes(x = mpg)) +
  geom_histogram(binwidth=1, aes(y=..density..), fill="#377EB8")


```
```{r}
# Draw a bar plot of cyl, filled according to am
ggplot(mtcars, aes(x = factor(cyl), fill = factor(am))) +
  geom_bar()

# Change the position argument to stack
ggplot(mtcars, aes(x = factor(cyl), fill = factor(am))) +
  geom_bar(position="stack")


# Change the position argument to fill
ggplot(mtcars, aes(x = factor(cyl), fill = factor(am))) +
  geom_bar(position="fill")


# Change the position argument to dodge
ggplot(mtcars, aes(x = factor(cyl), fill = factor(am))) +
  geom_bar(position="dodge")

```
Overlapping bar plots:
```{r}
# 1 - The last plot form the previous exercise
ggplot(mtcars, aes(x = factor(cyl), fill = factor(am))) +
  geom_bar(position = "dodge")

# 2 - Define posn_d with position_dodge()
posn_d <- position_dodge(width=0.2)

# 3 - Change the position argument to posn_d
ggplot(mtcars, aes(x = factor(cyl), fill = factor(am))) +
  geom_bar(position = posn_d)


# 4 - Use posn_d as position and adjust alpha to 0.6
ggplot(mtcars, aes(x = factor(cyl), fill = factor(am))) +
  geom_bar(position = posn_d, alpha=0.6)


```
Overlapping histograms
```{r}
# A basic histogram, add coloring defined by cyl
ggplot(mtcars, aes(mpg, fill=cyl)) +
  geom_histogram(binwidth = 1)

# Change position to identity
ggplot(mtcars, aes(mpg, fill=cyl)) +
  geom_histogram(binwidth = 1, position="identity")



# Change geom to freqpoly (position is identity by default)
ggplot(mtcars, aes(mpg, col=cyl)) +
  geom_freqpoly(binwidth=1)

```

Bar plots with color ramp
```{r}
# Example of how to use a brewed color palette
ggplot(mtcars, aes(x = factor(cyl), fill = factor(am))) +
  geom_bar() +
  scale_fill_brewer(palette = "Set1")

# Use str() on Vocab to check out the structure
str(Vocab)

# Plot education on x and vocabulary on fill
# Use the default brewed color palette
ggplot(Vocab, aes(factor(education), fill=factor(vocabulary)))+geom_bar(position="fill")+
scale_fill_brewer()


```
```{r}
library(RColorBrewer)

# Final plot of last exercise
ggplot(Vocab, aes(factor(education), fill=factor(vocabulary))) +
  geom_bar(position = "fill") +
  scale_fill_brewer()

# Definition of a set of blue colors
blues <- brewer.pal(9, "Blues") # from the RColorBrewer package

# 1 - Make a color range using colorRampPalette() and the set of blues
blue_range <- colorRampPalette(blues)

# 2 - Use blue_range to adjust the color of the bars, use scale_fill_manual()
scale_fill_manual(values=blue_range(11))
ggplot(Vocab, aes(factor(education), fill=factor(vocabulary))) +
  geom_bar(position = "fill") +
  scale_fill_manual(values=blue_range(11))


```
Overlapping histograms - Part 2
```{r}
# 1 - Basic histogram plot command
ggplot(mtcars, aes(mpg)) +
  geom_histogram(binwidth = 1)

# 2 - Plot 1, Expand aesthetics: am onto fill
ggplot(mtcars, aes(mpg, fill=factor(am))) +
  geom_histogram(binwidth = 1)


# 3 - Plot 2, change position = "dodge"
ggplot(mtcars, aes(mpg, fill=factor(am))) +
  geom_histogram(binwidth = 1, position="dodge")


# 4 - Plot 3, change position = "fill"
ggplot(mtcars, aes(mpg, fill=factor(am))) +
  geom_histogram(binwidth = 1, position="fill")


# 5 - Plot 4, plus change position = "identity" and alpha = 0.4
ggplot(mtcars, aes(mpg, fill=factor(am))) +
  geom_histogram(binwidth = 1, position="identity", alpha=0.4)


# 6 - Plot 5, plus change mapping: cyl onto fill
ggplot(mtcars, aes(mpg, fill=factor(am))) +
  geom_histogram(binwidth = 1, position="identity", alpha=0.4)

```
Line plots - Time series
```{r}
# Print out head of economics
head(economics)

# Plot unemploy as a function of date using a line plot
ggplot(economics, aes(x = date, y = unemploy)) + geom_line()

# Adjust plot to represent the fraction of total population that is unemployed
ggplot(economics, aes(x = date, y = unemploy/pop)) + geom_line()
```
```{r}
# Basic line plot
recess <-read.csv("Recess.csv")
recess$begin<- as.Date(recess$begin)
recess$end <- as.Date(recess$end)

ggplot(economics, aes(x = date, y = unemploy/pop)) +
  geom_line()

# Expand the following command with geom_rect() to draw the recess periods


```
Creating time series analysis using ggplot2
```{r}
fish.species<-read.csv("fish.species.csv")
# Check the structure as a starting point
str(fish.species)


# Use gather to go from fish.species to fish.tidy
fish.tidy <- gather(fish.species, Species, Capture, -Year)

# Recreate the plot shown on the right
ggplot(fish.tidy, aes(x = Year, y = Capture, col=Species)) + geom_line()

```
Qplot

simpler than ggplot is better for easy plots , less flexible than ggplot2
NOTES: There is no actual difference between the ggplot and qplot. Base plot defaults are slightly different. Plots 1 and 2 are base, 3 and 4 are ggplot
```{r}
#Basic Q plot
# The old way (shown)
plot(mpg ~ wt, data = mtcars) # formula notation
with(mtcars, plot(wt, mpg)) # x, y notation

# Using ggplot:
ggplot(mtcars, aes(wt, mpg)) +
  geom_point()

# Using qplot:
qplot(wt, mpg, data = mtcars)

```
Aesthetics in qplot
```{r}
# basic qplot scatter plot:
qplot(wt, mpg, data = mtcars)

# Categorical variable mapped onto size:
# cyl
qplot(wt, mpg, data = mtcars, size = factor(cyl))

# gear
qplot(wt, mpg, data = mtcars, size = factor(gear))

# Continuous variable mapped onto col:
# hp
qplot(wt, mpg, data = mtcars, col = hp)

# qsec
qplot(wt, mpg, data = mtcars, col = qsec)
```
Jitter inside the function call
```{r}
# qplot() with x only
qplot(factor(cyl), data=mtcars)

# qplot() with x and y
qplot(factor(cyl),factor(vs), data=mtcars)


# qplot() with geom set to jitter manually
qplot(factor(cyl), factor(vs), data=mtcars, geom="jitter")
```
Choosing geoms - part 2
```{r}
# cyl and am are factors, wt is numeric
class(mtcars$cyl)
class(mtcars$am)
class(mtcars$wt)

# "Basic" dot plot, with geom_point():
ggplot(mtcars, aes(factor(cyl), wt, col = factor(am))) +
  geom_point(position = position_jitter(0.2, 0))

# 1 - "True" dot plot, with geom_dotplot():
ggplot(mtcars, aes(factor(cyl), wt, fill = factor(am))) +
  geom_dotplot(binaxis = "y", stackdir = "center")

# 2 - qplot with geom "dotplot", binaxis = "y" and stackdir = "center"
qplot(
  factor(cyl), wt,
  data = mtcars,
  fill = factor(am),
  geom = "dotplot",
  binaxis = "y",
  stackdir = "center"
)
```
GGPlot advanced - Chickweight
```{r}
# ChickWeight is available in your workspace
# 1 - Check out the head of ChickWeight
head(ChickWeight)

# 2 - Basic line plot
ggplot(ChickWeight, aes(x = Time, y = weight)) + geom_line(aes(group = Chick))

# 3 - Take plot 2, map Diet onto col.
ggplot(ChickWeight, aes(x = Time, y = weight, col=Diet)) + geom_line(aes(group = Chick))


# 4 - Take plot 3, add geom_smooth()
ggplot(ChickWeight, aes(x = Time, y = weight, col=Diet)) + geom_line(aes(group = Chick),alpha=0.3)+geom_smooth(lwd=2,se=F)


```
Advanced visualization using ggplot2
```{r}
# titanic is avaliable in your workspace
titanic<-read.csv("titanic.csv")
# titanic is avaliable in your workspace
# 1 - Check the structure of titanic
str(titanic)

# 2 - Use ggplot() for the first instruction
ggplot(titanic, aes(x = Pclass, fill = Sex)) +
  geom_bar(position = "dodge")

# 3 - Plot 2, add facet_grid() layer
ggplot(titanic, aes(x = Pclass, fill = Sex)) +
  geom_bar(position = "dodge")+ facet_grid(.~ Survived)

# 4 - Define an object for position jitterdodge, to use below
posn.jd <- position_jitterdodge(0.5, 0, 0.6)

# 5 - Plot 3, but use the position object from instruction 4
ggplot(titanic, aes(x = Pclass,y=Age, col = Sex)) +
  geom_point(size=3, alpha=0.5, position = posn.jd)+ facet_grid(.~ Survived)



```

