---
title: "R Notebook"
output: html_notebook
---

Loading the libraries
```{r}
library(ggplot2)
library(RColorBrewer)
library(Hmisc)
library(GGally)
```

Statistics Layer
within a geom
independently

stats_ - stat fuctions

stat_smooth - standard error - gray ribbon with 95% confidence interval

span - alpha - degree of smoothing

methods - lm, glm, 

se=F -> ribbon is not visible

Playing around with geom_smooth
```{r}
# ggplot2 is already loaded

# Explore the mtcars data frame with str()
str(mtcars)

# A scatter plot with LOESS smooth
ggplot(mtcars, aes(x = wt, y = mpg)) +
  geom_point()


# A scatter plot with an ordinary Least Squares linear model
ggplot(mtcars, aes(x = wt, y = mpg)) +
  geom_point() + geom_smooth(method="lm")

# The previous plot, without CI ribbon
ggplot(mtcars, aes(x = wt, y = mpg)) + geom_point()+
  geom_smooth(method="lm", se=F)


# The previous plot, without points
ggplot(mtcars, aes(x = wt, y = mpg)) + geom_smooth(method="lm", se=F)

```
Stat smooth groupings
```{r}
# ggplot2 is already loaded

# 1 - Define cyl as a factor variable
ggplot(mtcars, aes(x = wt, y = mpg, col = factor(cyl))) +
  geom_point() +
  stat_smooth(method = "lm", se = FALSE)

# 2 - Plot 1, plus another stat_smooth() containing a nested aes()
ggplot(mtcars, aes(x = wt, y = mpg, col = factor(cyl))) +
  geom_point() +
  stat_smooth(method = "lm", se = FALSE) +
  stat_smooth(method = "lm", se = FALSE, aes(group=1))



```
Advanced smoothening
```{r}
# Plot 1: change the LOESS span
ggplot(mtcars, aes(x = wt, y = mpg)) +
  geom_point() +
  # Add span below
  geom_smooth(se = FALSE, span=0.7)

# Plot 2: Set the second stat_smooth() to use LOESS with a span of 0.7
ggplot(mtcars, aes(x = wt, y = mpg, col = factor(cyl))) +
  geom_point() +
  stat_smooth(method = "lm", se = FALSE) +
  # Change method and add span below
  stat_smooth(method = "loess", aes(group = 1),
              se = FALSE, col = "black", span=0.7)

# Plot 3: Set col to "All", inside the aes layer of stat_smooth()
ggplot(mtcars, aes(x = wt, y = mpg, col = factor(cyl))) +
  geom_point() +
  stat_smooth(method = "lm", se = FALSE) +
  stat_smooth(method = "loess",
              # Add col inside aes()
              aes(group = 1, col="All"),
              # Remove the col argument below
              se = FALSE, span = 0.7)

# Plot 4: Add scale_color_manual to change the colors
myColors <- c(brewer.pal(3, "Dark2"), "black")
ggplot(mtcars, aes(x = wt, y = mpg, col = factor(cyl))) +
  geom_point() +
  stat_smooth(method = "lm", se = FALSE, span = 0.7) +
  stat_smooth(method = "loess", 
              aes(group = 1, col="All"), 
              se = FALSE, span = 0.7) +
  # Add correct arguments to scale_color_manual
  scale_color_manual("Cylinders", values=myColors)
```
```{r}
Vocab<-read.csv("Vocab.csv")
# Plot 1: Jittered scatter plot, add a linear model (lm) smooth
ggplot(Vocab, aes(x = education, y = vocabulary)) +
  geom_jitter(alpha = 0.2) +
  stat_smooth(method="lm", se=F) # smooth

# Plot 2: points, colored by year
ggplot(Vocab, aes(x = education, y = vocabulary, col = year)) +
  geom_jitter(alpha = 0.2) 

# Plot 3: lm, colored by year
ggplot(Vocab, aes(x = education, y = vocabulary, col = factor(year))) +
  stat_smooth(method="lm", se=F) # smooth
  
# Plot 4: Set a color brewer palette
ggplot(Vocab, aes(x = education, y = vocabulary, col = factor(year))) +
  stat_smooth(method="lm", se=F) +  # smooth
  scale_color_brewer()  # colors

# Plot 5: Add the group aes, specify alpha and size
ggplot(Vocab, aes(x = education, y = vocabulary, col = year, group = factor(year))) +
  stat_smooth(method = "lm", se = FALSE, alpha=0.6, size=2) +
  scale_color_gradientn(colors = brewer.pal(9, "YlOrRd"))
```
```{r}
options(warn = 0)
# Use stat_quantile instead of stat_smooth
ggplot(Vocab, aes(x = education, y = vocabulary, col = year, group = factor(year))) +
  stat_quantile(alpha = 0.6, size = 2) +
  scale_color_gradientn(colors = brewer.pal(9,"YlOrRd"))

# Set quantile to 0.5
ggplot(Vocab, aes(x = education, y = vocabulary, col = year, group = factor(year))) +
  stat_quantile(quantiles=0.5 , alpha = 0.6, size = 2) +
  scale_color_gradientn(colors = brewer.pal(9,"YlOrRd"))


```

```{r}
# Plot 1: Jittering only
p <- ggplot(Vocab, aes(x = education, y = vocabulary)) +
  geom_jitter(alpha = 0.2)

# Plot 2: Add stat_sum
p + stat_sum()# sum statistic

# Plot 3: Set size range
p + stat_sum()+# sum statistic
  scale_size(range = c(1, 10)) # set size scale

```
```{r}
# Display structure of mtcars
str(mtcars)

# Convert cyl and am to factors
mtcars$cyl <- factor(mtcars$cyl)
mtcars$am <- factor(mtcars$am)

# Define positions
posn.d <- position_dodge(width=0.1)
posn.jd <- position_jitterdodge(jitter.width=0.1, dodge.width=0.2) 
posn.j <- position_jitter(width=0.2)

# Base layers
wt.cyl.am <- ggplot(mtcars, aes(cyl, wt, group=am, colour=am, fill=am))
## Advanced statistics using base layers and advanced fundamentals
# wt.cyl.am, posn.d, posn.jd and posn.j are available

# Plot 1: Jittered, dodged scatter plot with transparent points
wt.cyl.am +
  geom_point(position = posn.jd, alpha = 0.6)

# Plot 2: Mean and SD - the easy way
wt.cyl.am + stat_summary(fun.data=mean_sdl, fun.args=list(mult=1), position=posn.d)


# Plot 3: Mean and 95% CI - the easy way
wt.cyl.am + stat_summary(fun.data=mean_cl_normal, position=posn.d)



# Plot 4: Mean and SD - with T-tipped error bars - fill in ___
wt.cyl.am +
  stat_summary(geom = "point", fun.y = mean,
               position = posn.d) +
  stat_summary(geom = "errorbar", fun.data = mean_sdl,
               position = posn.d, fun.args = list(mult = 1), width = 0.1)
```
Advanced statistics in R
```{r}

```

```{r}
##Getting quantiles of variable
xx <- c(1,2,3,4,5,6,7,8,9,10,11,12,13,14,15,16,17,18,19,20,21,22,23,24,25,26,27,28,29,30,31,32,33,34,35,36,37,38,39,40,41,42,43,44,45,46,47,48,49,50,51,52,53,54,55,56,57,58,59,60,61,62,63,64,65,66,67,68,69,70,71,72,73,74,75,76,77,78,79,80,81,82,83,84,85,86,87,88,89,90,91,92,93,94,95,96,97,98,99,100)
xx
# Play vector xx is available
xx
# Function to save range for use in ggplot
gg_range <- function(x) {
  # Change x below to return the instructed values
  data.frame(ymin = min(x), # Min
             ymax = max(x)) # Max
}

gg_range(xx)
# Required output
#   ymin ymax
# 1    1  100

# Function to Custom function
med_IQR <- function(x) {
  # Change x below to return the instructed values
  data.frame(y = median(x), # Median
             ymin = quantile(x)[2], # 1st quartile
             ymax = quantile(x)[4]  )# 3rd quartile
}

med_IQR(xx)
# Required output
#        y  ymin  ymax
# 25% 50.5 25.75 75.25
```
Making custom plots in r
```{r}
# The base ggplot command; you don't have to change this
wt.cyl.am <- ggplot(mtcars, aes(x = cyl,y = wt, col = am, fill = am, group = am))

# Add three stat_summary calls to wt.cyl.am
wt.cyl.am +
  stat_summary(geom = "linerange", fun.data = med_IQR,
               position = posn.d, size = 3) +
  stat_summary(geom = "linerange", fun.data = gg_range,
               position = posn.d, size = 3,
               alpha = 0.4) +
  stat_summary(geom = "point", fun.y = median,
               position = posn.d, size = 3,
               col = "black", shape = "X")
```
Coordinates layer
Controls plot dimensions
  coord_  
  coord_cartesian - controls xy plane

Zooming in
  -scale_x_continuous(lim)
  -xlim
  -coords_cartisian - pure zooming in
  
Aspect ratio
  Generally 1:1 when x and y are same
  Sublte inferences can be observed using extreme aspect ratio settings
  
Part 1 - Zooming in
```{r}
# Basic ggplot() command, coded for you
p <- ggplot(mtcars, aes(x = wt, y = hp, col = am)) + geom_point() + geom_smooth()

# Add scale_x_continuous()
p + scale_x_continuous(limits=c(3,6), expand=c(0,0))

# Add coord_cartesian(): the proper way to zoom in
p + coord_cartesian(xlim=c(3,6))

```
Aspect ratios - 
```{r}
# Complete basic scatter plot function
base.plot <- ggplot(iris, aes(x = Sepal.Length, y = Sepal.Width, col = Species)) +
               geom_jitter() +
               geom_smooth(method = "lm", se = FALSE)

# Plot base.plot: default aspect ratio
base.plot

# Fix aspect ratio (1:1) of base.plot
base.plot + coord_equal()
```

```{r}
# Create a stacked bar plot: wide.bar
wide.bar <- ggplot(mtcars, aes(x = 1, fill = cyl)) +
              geom_bar()

# Convert wide.bar to pie chart
wide.bar +
  coord_polar(theta = "y")

# Create stacked bar plot: thin.bar
thin.bar <- ggplot(mtcars, aes(x=1, fill = cyl)) +
              geom_bar(width = 0.1) +
              scale_x_continuous(limits = c(0.5,1.5))

# Convert thin.bar to "ring" type pie chart
thin.bar + 
  coord_polar(theta="y")
```
Facets - Small multiples - Each has the same coordinates so that they can be easily copared
A ~ B
A Rows
B Columns
```{r}
# Basic scatter plot
p <- ggplot(mtcars, aes(x = wt, y = mpg)) +
  geom_point()
p
# 1 - Separate rows according to transmission type, am
p +
  facet_grid(am ~ .)

# 2 - Separate columns according to cylinders, cyl
p +
  facet_grid(. ~ cyl)

# 3 - Separate by both columns and rows 
p +
  facet_grid(am ~ cyl)

```

```{r}
# Code to create the cyl_am col and myCol vector
mtcars$cyl_am <- paste(mtcars$cyl, mtcars$am, sep = "_")
myCol <- rbind(brewer.pal(9, "Blues")[c(3,6,8)],
               brewer.pal(9, "Reds")[c(3,6,8)])

# Map cyl_am onto col
ggplot(mtcars, aes(x = wt, y = mpg, col = cyl_am)) +
  geom_point() +
  # Add a manual colour scale
  scale_color_manual(values = myCol)

  
# Grid facet on gear vs. vs
ggplot(mtcars, aes(x = wt, y = mpg, col = cyl_am)) +
  geom_point() +
  # Add a manual colour scale
  scale_color_manual(values = myCol)
  




# Also map disp to size
ggplot(mtcars, aes(x = wt, y = mpg, col = cyl_am, size=disp)) +
  geom_point() +
  # Add a manual colour scale
  scale_color_manual(values = myCol)
  



```
```{r}
### REFER TO CHAPTER FOR EXECUTION - DROPPING LEVELS in facets
# Basic scatter plot
p <- ggplot(mamsleep, aes(x = time, y = name, col = sleep)) +
  geom_point()
  
# Execute to display plot
p

# Facet rows accoding to vore
p +
  facet_grid(vore ~ .)

# Specify scale and space arguments to free up rows
p +
  facet_grid(vore ~ ., scale="free_y", space="free_y")

```

Themes - Non data ink
 theme(
 text - all text elements
 line - all line elements 
 rectangle - all rectangle elements
 )
 
 element_blank() - element is not plotted

```{r}
myColBlue <- brewer.pal(9, "Blues")[c(3,6,8)]
myPink <-"#FEE0D2"
               
 z <- ggplot(mtcars, aes(wt,mpg, col=factor(cyl)))+geom_point() + stat_smooth(method = "lm", se = FALSE) + scale_color_manual(values = myColBlue)+facet_grid(.~factor(cyl))
# Starting point
z

# Plot 1: Change the plot background fill to myPink
z +
  theme(plot.background = element_rect(fill = myPink))

# Plot 2: Adjust the border to be a black line of size 3
z +
  theme(plot.background = element_rect(fill = myPink, color="black", size=3)) # expanded from plot 1

# Theme to remove all rectangles
no_panels <- theme(rect = element_blank())

# Plot 3: Combine custom themes
z +
  no_panels +
  theme(plot.background = element_rect(fill = myPink, color="black", size=3))# from plot 2

# Extend z using theme() function and 3 args
z+theme(panel.grid=element_blank(), axis.line=element_line(color="red"), axis.ticks=element_line(color="red"))
```

```{r}
# Original plot, color provided
z
myRed

# Extend z with theme() function and 3 args
z +
  theme(strip.text = element_text(size =16, color = myRed),
        axis.title = element_text(color = myRed, hjust = 0, face = "italic"),
        axis.text = element_text(color = "Black"))
```
```{r}
# Move legend by position
z +
  theme(legend.position = c(0.85,0.85))

# Change direction
z +
  theme(legend.direction= "horizontal")
  
# Change location by name
z +
  theme(legend.position = "bottom")

# Remove legend entirely
z +
  theme(legend.position = "none")
```
Facet grid margins
```{r}
# Increase spacing between facets
library(grid)
z+theme(panel.spacing.x=unit(2,"cm"))


# Adjust the plot margin
z+theme(panel.spacing.x=unit(2,"cm"), plot.margin=unit(c(1,2,1,1), "cm"))
```

Recycling themes
- Many plots should have consistency in style
Save theme function as an object that can be used over different plots

Using built-in theme templates
ggthemes packages - tufte theme
theme update function - Save 

theme_set - will save default theme for all plots
```{r}
# Original plot
z2

# Theme layer saved as an object, theme_pink
theme_pink <- theme(panel.background = element_blank(),
                    legend.key = element_blank(),
                    legend.background = element_blank(),
                    strip.background = element_blank(),
                    plot.background = element_rect(fill = myPink, color = "black", size = 3),
                    panel.grid = element_blank(),
                    axis.line = element_line(color = "red"),
                    axis.ticks = element_line(color = "red"),
                    strip.text = element_text(size = 16, color = myRed),
                    axis.title.y = element_text(color = myRed, hjust = 0, face = "italic"),
                    axis.title.x = element_text(color = myRed, hjust = 0, face = "italic"),
                    axis.text = element_text(color = "black"),
                    legend.position = "none")
  
# 1 - Apply theme_pink to z2
z2 + theme_pink

# 2 - Update the default theme, and at the same time
# assign the old theme to the object old.
old <- theme_update(panel.background = element_blank(),
             legend.key = element_blank(),
             legend.background = element_blank(),
             strip.background = element_blank(),
             plot.background = element_rect(fill = myPink, color = "black", size = 3),
             panel.grid = element_blank(),
             axis.line = element_line(color = "red"),
             axis.ticks = element_line(color = "red"),
             strip.text = element_text(size = 16, color = myRed),
             axis.title.y = element_text(color = myRed, hjust = 0, face = "italic"),
             axis.title.x = element_text(color = myRed, hjust = 0, face = "italic"),
             axis.text = element_text(color = "black"),
             legend.position = "none")

# 3 - Display the plot z2 - new default theme used
z2

# 4 - Restore the old default theme
theme_set(old)

# Display the plot z2 - old theme restored
z2

```
Custom theme
```{r}
# Original plot
z2

# Load ggthemes
library(ggthemes)

# Apply theme_tufte(), plot additional modifications
custom_theme <- theme_tufte()+
  theme(legend.position = c(0.9,0.9),
        legend.title = element_text(size=12, face="italic"),
        axis.title = element_text(face="bold", size=14))
# Draw the customized plot
z2 + custom_theme
 
# Use theme set to set custom theme as default
theme_set(custom_theme)

# Plot z2 again
z2
```

Data viz best practices

Common pitfalls
Dynamite plots - careful to use

Scatter plots are better than bar plots
points are better than bars

```{r}
# Base layers
m <- ggplot(mtcars, aes(x = factor(cyl), y = wt))

# Draw dynamite plot
m +
  stat_summary(fun.y = mean, geom = "bar", fill = "skyblue") +
  stat_summary(fun.data = mean_sdl, fun.args = list(mult = 1), geom = "errorbar", width = 0.1)
```
```{r}
# Base layers
m <- ggplot(mtcars, aes(x = factor(cyl),y = wt, col = factor(am), fill = factor(am)))

# Plot 1: Draw dynamite plot
m +
  stat_summary(fun.y = mean, geom = "bar") +
  stat_summary(fun.data = mean_sdl, fun.args = list(mult = 1), geom = "errorbar", width = 0.1)

# Plot 2: Set position dodge in each stat function
m +
  stat_summary(fun.y = mean, geom = "bar", position = "dodge") +
  stat_summary(fun.data = mean_sdl, fun.args = list(mult = 1), 
               geom = "errorbar", width = 0.1, position = "dodge")

# Set your dodge posn manually
posn.d <- position_dodge(0.9)

# Plot 3: Redraw dynamite plot
m +
  stat_summary(fun.y = mean, geom = "bar", position = posn.d) +
  stat_summary(fun.data = mean_sdl, fun.args = list(mult = 1), geom = "errorbar", width = 0.1, position = posn.d)
```

```{r}
mtcars.cyl<-read.csv("Mtcarscyl.csv")
mtcars.cyl
# Base layers
m <- ggplot(mtcars.cyl, aes(x = cyl, y = wt.avg))

# Plot 1: Draw bar plot with geom_bar
m + geom_bar(stat = "identity", fill = "skyblue")

# Plot 2: Draw bar plot with geom_col
m + geom_col(fill = "skyblue")

# Plot 3: geom_col with variable widths.
m + geom_col(fill = "skyblue", width = mtcars.cyl$prop)
 
# Plot 4: Add error bars
m + 
  geom_col(fill = "skyblue", width = mtcars.cyl$prop)
  geom_errorbar(aes(ymin = wt.avg-sd, ymax = wt.avg+sd), width = 0.1)
```

Pie charts - Why are they not important
Parts of whole?
Can all categories be clearly represented?
Good - Large data in max 3 groups
For other things, use stacked bar plot, parallel plots

```{r}
# Bar chart
ggplot(mtcars, aes(x = factor(cyl), fill = factor(am))) +
  geom_bar(position = "fill")

# Convert bar chart to pie chart
ggplot(mtcars, aes(x = factor(1), fill = factor(am))) +
  geom_bar(position = "fill", width=1) +
  facet_grid(. ~ cyl) + # Facets
  coord_polar(theta = "y") + # Coordinates
  theme_void() # theme
  
```
```{r}
# Parallel coordinates plot using GGally
library(GGally)

# All columns except am
group_by_am <- 9
my_names_am <- (1:11)[-group_by_am]

# Basic parallel plot - each variable plotted as a z-score transformation
ggparcoord(mtcars, my_names_am, groupColumn = group_by_am, alpha = 0.8)
```
Heat maps
Generally ineffective
Violates tufte's principles
Mapping color to continuous is not good
Not good - Perception of color changes
Good for specific questions

Alt - Plotting data on commong scale is better option

```{r}
# Create color palette
myColors <- brewer.pal(9, "Reds")

# Build the heat map from scratch
ggplot(barley, aes(x = year, y = variety, fill = yield)) +
  geom_tile() + # Geom layer
  facet_wrap( ~ site, ncol = 1) + # Facet layer
  scale_fill_gradientn(colors = myColors) # Adjust colors
```
Heat map alternatives
```{r}
# The heat map we want to replace
# Don't remove, it's here to help you!
myColors <- brewer.pal(9, "Reds")
ggplot(barley, aes(x = year, y = variety, fill = yield)) +
  geom_tile() +
  facet_wrap( ~ site, ncol = 1) +
  scale_fill_gradientn(colors = myColors)

# Line plot; set the aes, geom and facet
ggplot(barley, aes(x = year, y = yield, col = variety, group=variety)) + geom_line()+ facet_wrap(~ site, nrow =1 )

```
```{r}
# Create overlapping ribbon plot from scratch
ggplot(barley, aes(x=year, y=yield, col=site, group = site, fill= site)) + stat_summary(fun.y=mean, geom="line")+ stat_summary(fun.data = mean_sdl, fun.args = list(mult =1), geom="ribbon", col=NA, alpha=0.1)
```

